---
title: "Preprocessing rVTS 2018"
author: "Dave Braun"
date: "May 7, 2022"
output:
  html_document:
    code_folding: hide
    theme: flatly
    df_print: paged
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
    encoding = encoding,
    output_file = 'index.html')})
  
---


This document was last updated at `r Sys.time()`.

In one shot I'm going to clean the foraging and disconnect data.  

* Subjects  
  * Drop if errors higher than 15%  
* Trials  
  * Drop all RTs longer than 10s  
  * Drop first trial of block  
  * Drop error trials and trials following error trials  
  * Trim on +\- 2 SDs from subject-wise RT means
  
  
## Import and combine

```{r}
library(tidyverse)
library(data.table)
```


```{r}
## switching to processing one dataset at a time
cued <- read.csv('../../../data/raw_data/foraging_cued_full.csv')
rvts <- read.csv('../../../data/raw_data/foraging_rvts_full.csv')

cued_keep <- c('workerId', 'rt', 'transition', 'trial', 'error')
rvts_keep <- c('workerId', 'blockTime', 'Block', 'rt', 'totalPoints', 'transition', 
               'rightpoints', 'leftpoints', 'response_location', 'error', 'trial')

rvts <- rvts[, rvts_keep]
cued <- cued[, cued_keep]
```


```{r}
## compute vars
cued$error_trim <- ifelse(cued$error == 1 | shift(cued$error == 1), 1, 0)
rvts$error_trim <- ifelse(rvts$error == 1 | shift(rvts$error == 1), 1, 0)

cued$transcode <- ifelse(cued$transition == 'Switch', 1, 0)
rvts$transcode <- ifelse(rvts$transition == 'Switch', 1, 0)

colnames(cued)[colnames(cued) == 'workerId'] <- 'subject'
colnames(rvts)[colnames(rvts) == 'workerId'] <- 'subject'

rvts$difference <- ifelse(shift(rvts$response_location) == 'left', 
                          rvts$rightpoints - rvts$leftpoints, rvts$leftpoints - rvts$rightpoints)
rvts <- rvts %>% 
  mutate(point_transition = ifelse(shift(rightpoints) == 1 & rightpoints == 3, 'increasing',
                                   ifelse(shift(rightpoints) == 3 & rightpoints == 3, 'stay high',
                                   ifelse(shift(rightpoints) == 1 & rightpoints == 1, 'stay low',
                                   ifelse(shift(rightpoints) == 3 & rightpoints == 1, 'decreasing', '')))))

rvts$current_points <- ifelse(shift(rvts$response_location) == 'left', rvts$leftpoints, ifelse(shift(rvts$response_location) == 'right', rvts$rightpoints, NA))

rvts$other_points <- ifelse(shift(rvts$response_location) == 'right', rvts$leftpoints, ifelse(shift(rvts$response_location) == 'left', rvts$rightpoints, NA))

rvts$current_location <- ifelse(shift(rvts$response_location) == 'left', 'left', 'right')
rvts$other_location <- ifelse(shift(rvts$response_location) == 'left', 'right', 'left')

rvts$left_transition <- ifelse(rvts$leftpoints < shift(rvts$leftpoints), 'decrease',
                               ifelse(rvts$leftpoints == shift(rvts$leftpoints), 'constant',
                                      ifelse(rvts$leftpoints > shift(rvts$leftpoints), 'increase', NA)))

rvts$right_transition <- ifelse(rvts$rightpoints < shift(rvts$rightpoints), 'decrease',
                               ifelse(rvts$rightpoints == shift(rvts$rightpoints), 'constant',
                                      ifelse(rvts$rightpoints > shift(rvts$rightpoints), 'increase', NA)))


rvts$current_transition <- ifelse(rvts$current_location == 'left', rvts$left_transition, ifelse(rvts$current_location == 'right', rvts$right_transition, NA))

rvts$other_transition <- ifelse(rvts$other_location == 'left', rvts$left_transition, ifelse(rvts$other_location == 'right', rvts$right_transition, NA))


```

```{r}
## drop subjects where i forgot to adjust the point threshold 
pilot_subjects <- rvts %>% 
  group_by(subject) %>% 
  summarize(max_points = max(totalPoints)) %>% 
  filter(max_points < 100) %>% 
  select(subject)

rvts <- rvts[!rvts$subject %in% pilot_subjects$subject,]
cued <- cued[!cued$subject %in% pilot_subjects$subject,]
initialN <- length(unique(rvts$subject))
```

There were **`r initialN` initial subjects.**

## Trim data

### Trim subjects

```{r}
bad_subjects <- rvts %>% 
  group_by(subject) %>% 
  summarize(error = mean(error)) %>% 
  filter(error > .15) %>% 
  select(subject)

rvts <- rvts[!rvts$subject %in% bad_subjects$subject,]
cued <- cued[!cued$subject %in% bad_subjects$subject,]
```



For the cued data, all I want is the btwn subject switch costs, so I'm going to compute those only after dropping > 10s RTs.
```{r}
trim_summary <- data.frame('reason' = c(), 'proportion_dropped' = c(), rows_dropped = c(), total_rows = c(), phase = c())

start <- nrow(cued)
cued <- cued[cued$rt < 10000,]
trim_summary <- rbind(trim_summary, data.frame('reason' = 'RTs over 10s', 
                                               'proportion_dropped' = 1 - (nrow(cued) / start), phase = 'cued', rows_dropped = start - nrow(cued), total_rows = start))
start <- nrow(cued)
cued <- cued[cued$transition != 'StartBlock',]
trim_summary <- rbind(trim_summary, data.frame('reason' = 'StartBlock', 
                                               'proportion_dropped' = 1 - (nrow(cued) / start), phase = 'cued', rows_dropped = start - nrow(cued), total_rows = start))

```

```{r}
## compute switch costs
switch_costs <- cued %>% 
  group_by(subject, transition) %>% 
  summarize(rt = mean(rt)) %>% 
  spread(transition, rt) %>% 
  mutate(switch_cost = Switch - Repeat) %>% 
  select(-Repeat, -Switch) 

hist(switch_costs$switch_cost, main = 'Switch Costs')
```

## Trim the rest of rVTS data

```{r}
start <- nrow(rvts)
rvts <- rvts[rvts$transition != 'StartBlock',]
trim_summary <- rbind(trim_summary, data.frame('reason' = 'Startblock', 
                                               'proportion_dropped' = 1 - (nrow(rvts) / start), phase = 'rvts', rows_dropped = start - nrow(rvts), total_rows = start))

start <- nrow(rvts)
rvts <- rvts[rvts$rt < 10000,]
trim_summary <- rbind(trim_summary, data.frame('reason' = 'RTs longer than 10s', 
                                               'proportion_dropped' = 1 - (nrow(rvts) / start), phase = 'rvts', rows_dropped = start - nrow(rvts), total_rows = start))

start <- nrow(rvts)
rvts <- rvts %>% 
  group_by(subject) %>% 
  summarize(rt_m = mean(rt), rt_sd = mean(rt)) %>% 
  inner_join(rvts) %>% 
  filter(rt <= rt_m + 2*rt_sd & rt >= rt_m - 2*rt_sd) %>% 
  select(-rt_m, -rt_sd)
trim_summary <- rbind(trim_summary, data.frame('reason' = '+/- 2 SDs on participant RT means', 
                                               'proportion_dropped' = 1 - (nrow(rvts) / start), phase = 'rvts', rows_dropped = start - nrow(rvts), total_rows = start))

rvts_errors <- rvts

start <- nrow(rvts)
rvts <- rvts[rvts$error_trim == 0, ]
trim_summary <- rbind(trim_summary, data.frame('reason' = 'Error and trials after errors', 
                                               'proportion_dropped' = 1 - (nrow(rvts) / start), phase = 'rvts', rows_dropped = start - nrow(rvts), total_rows = start))

```
## Trim cued data

```{r}
start <- nrow(cued)
cued <- cued[cued$rt < 10000,]
trim_summary <- rbind(trim_summary, data.frame('reason' = 'RTs longer than 10s', 
                                               'proportion_dropped' = 1 - (nrow(cued) / start), phase = 'cued', rows_dropped = start - nrow(cued), total_rows = start))

start <- nrow(cued)
cued <- cued %>% 
  group_by(subject) %>% 
  summarize(rt_m = mean(rt), rt_sd = mean(rt)) %>% 
  inner_join(cued) %>% 
  filter(rt <= rt_m + 2*rt_sd & rt >= rt_m - 2*rt_sd) %>% 
  select(-rt_m, -rt_sd)
trim_summary <- rbind(trim_summary, data.frame('reason' = '+/- 2 SDs on participant RT means', 
                                               'proportion_dropped' = 1 - (nrow(cued) / start), phase = 'cued', rows_dropped = start - nrow(cued), total_rows = start))

cued_errors <- cued

start <- nrow(cued)
cued <- cued[cued$error_trim == 0, ]
trim_summary <- rbind(trim_summary, data.frame('reason' = 'Error and trials after errors', 
                                               'proportion_dropped' = 1 - (nrow(cued) / start), phase = 'cued', rows_dropped = start - nrow(cued), total_rows = start))

```



## Trimming summary

```{r}
bad_subjects
```
`r nrow(bad_subjects)` were dropped for having error rates over 15%.

```{r}
## add switch costs to rvts
rvts <- rvts %>% 
  inner_join(switch_costs) #%>% 
  #select(-transition, -rightpoints, -leftpoints, -response_location, -error_trim)
```
```{r}
## block times
rvts %>% 
  group_by(subject) %>% 
  summarize(block_time = mean(blockTime)) %>% 
  ggplot(aes(x = block_time)) + 
  geom_histogram(color = 'black')
```



```{r}
trim_summary
```

Final sample size was `r length(unique(rvts$subject))`

```{r}
write.csv(rvts, '../../../data/rvts_foraging.csv', row.names = FALSE)
write.csv(cued, '../../../data/cued_foraging.csv', row.names = FALSE)
write.csv(rvts_errors, '../../../data/rvts_errors_foraging.csv', row.names = FALSE)
write.csv(cued_errors, '../../../data/cued_errors_foraging.csv', row.names = FALSE)
```











