---
title: "Choice Analysis Foraging"
author: "Dave Braun"
date: "June 6, 2022"
output:
  html_document:
    code_folding: hide
    theme: flatly
    df_print: paged
    toc: true
    toc_float: true
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
    encoding = encoding,
    output_file = 'index.html')})
  
---


This document was last updated at `r Sys.time()`.

In this document I will perform all analyses related to task selection.

```{r include = FALSE}
library(tidyverse)
library(ez)
library(lme4)

d <- read.csv('../../../data/rvts_foraging.csv')
```

```{r}
N <- length(unique(d$subject))
head(d)
```

# Current & Other
  
```{r}
a <- d %>% 
  group_by(subject, current_transition, other_transition) %>% 
  summarize(sw_ = mean(transcode)) 
a %>% 
  group_by(current_transition, other_transition) %>% 
  summarize(sw = mean(sw_), se = sd(sw_) / sqrt(N)) %>% 
  mutate(current_transition = recode(current_transition, `decrease` = 'Current Value Decrease', `constant` = 'Current Value Constant'),
         other_transition = recode(other_transition, `increase` = 'Other Value Increase', `constant` = 'Other Value Constant')) %>% 
  ggplot(aes(x = current_transition, y = sw, group = other_transition)) +
  geom_bar(stat = 'identity', aes(fill = other_transition), position = position_dodge(width = .9), color = 'black') + 
  geom_errorbar(aes(ymin = sw - se, ymax = sw + se), position = position_dodge(width = .9), width = .3)+ 
  labs(
    x = '',
    y = 'Proportion of Task Switches',
    fill = ''
  ) + 
  ylim(0,1) + 
  scale_fill_manual(values = c('Other Value Constant' = 'dark grey', 'Other Value Increase' = 'white')) + 
  theme_bw() + 
  theme(axis.ticks = element_blank(),
        panel.grid = element_blank(),
        legend.position = c(.2, .8))
```


```{r}
m1 <- ezANOVA(wid = subject, within = .(current_transition, other_transition), dv = transcode, detailed = TRUE, data = d)
cbind(m1$ANOVA, data.frame(n2p = m1$ANOVA$SSn / (m1$ANOVA$SSn + m1$ANOVA$SSd)))

dfd <- 84
SSd <- m1$ANOVA$SSd[4]

im <- ezANOVA(wid = subject, within = .(other_transition), dv = transcode, detailed = TRUE, data = d[d$current_transition == 'constant',])
dfn <- 1
SSn <- im$ANOVA$SSn[2]

msb <- SSn / dfn
mse <- SSd / dfd

f <- msb / mse
p <- 1 - pf(f, dfn, dfd)
n2p <- SSn / (SSn + SSd)

print('Simple effect of other for current constant')
print(paste0('F(', dfn, ', ', dfd, ') = ', round(f, 2), ', ', round(p, 3), ', eta = ', round(n2p, 2)))
```

```{r}
im <- ezANOVA(wid = subject, within = .(other_transition), dv = transcode, detailed = TRUE, data = d[d$current_transition == 'decrease',])
dfn <- 1
SSn <- im$ANOVA$SSn[2]

msb <- SSn / dfn
mse <- SSd / dfd

f <- msb / mse
p <- 1 - pf(f, dfn, dfd)
n2p <- SSn / (SSn + SSd)

print('Simple effect of other for current decrease')
print(paste0('F(', dfn, ', ', dfd, ') = ', round(f, 2), ', ', round(p, 3), ', eta = ', round(n2p, 2)))
```


# Difference & Block -> PSwitch

```{r}
## code and center model vars
colnames(d) <- tolower(colnames(d))
## compute an impromptu trial count

d <- d %>% 
  group_by(subject) %>% 
  mutate(consec_trial = seq(n()))

d$consec_trial_sc <- scale(d$consec_trial)
```


```{r eval = FALSE}
## model fits
m2 <- glmer(transcode ~ consec_trial_sc * difference + (1 + difference | subject), data = d, control = glmerControl(optimizer = 'bobyqa'), family = binomial(link = 'logit'), nAGQ = 1)
m2_down1 <- glmer(transcode ~ consec_trial_sc * difference + (0 + difference | subject) + (1 | subject), data = d, control = glmerControl(optimizer = 'bobyqa'), family = binomial(link = 'logit'), nAGQ = 1)
m2_down2 <- glmer(transcode ~ consec_trial_sc * difference + (1 | subject), data = d, control = glmerControl(optimizer = 'bobyqa'), family = binomial(link = 'logit'), nAGQ = 1)
anova(m2_down2, m2_down1)
m2 <- m2_down1
save(m2, file = 'winning_glmm')
```

Winning model has random intercept and slope of difference (no cov).

Note: direction of interaction effect is opposite B&A2018. Other effects are consistent.

```{r}
load('winning_glmm')
summary(m2)
exp(fixef(m2))
exp(confint(m2, method = 'Wald'))
```


```{r}
m2_noint <- glmer(transcode ~ consec_trial_sc + difference + (0 + difference | subject) + (1 | subject), data = d, control = glmerControl(optimizer = 'bobyqa'), family = binomial(link = 'logit'), nAGQ = 1)
anova(m2, m2_noint)
```


*Still to do:* get chi square tests for both main effects. Decide what of these predictions I want to visualize, go get the `easypredci` function.























